---
title: Joint variant calling
author: ''
date: '2019-03-06'
slug: joint-variant-calling
categories:
  - Bioinformatics
tags:
  - GATK4
  - SNP
  - RNA-seq
---
When I tried to implement the pipeline for variant calling from RNA-seq data, I was confused about how to deal with biological replicated. According to https://gatkforums.broadinstitute.org/gatk/discussion/5894/variant-calling-with-biological-replicates , it seems that there are multiple ways to do this.

I tested three different ways on three replicates from HCT116b. In specific,the main pipeline(https://htmlpreview.github.io/?https://github.com/qibaotu/Cell-line-authentication/blob/master/vignette/Cell_line.html) is the same, except where to combine replicates.

I tested three different ways on three replicates from HCT116b. In specific,the main pipeline(https://htmlpreview.github.io/?https://github.com/qibaotu/Cell-line-authentication/blob/master/vignette/Cell_line.html)is the same, except where to combine replicates.


*Way1*. Combine bam file after `SplitNCigarReads`. In this way I would have one combined bam file to feed the `BaseRecalibrator`.
```{r,engine='bash',eval=FALSE, way1}
# merge bam file using samtools
samtools merge ./way1/HCT116b_dedup_split.bam SRR2481145_dedup_split.bam SRR2481146_dedup_split.bam SRR2481147_dedup_split.bam

## feed combined bam to the following step
gatk BaseRecalibrator \
        -R $genomeFastaFiles \
        -I HCT116b_dedup_split.bam \
        --known-sites 1000G_phase1.snps.high_confidence.hg38.vcf.gz \
        --known-sites Mills_and_1000G_gold_standard.indels.hg38.vcf.gz \
		    --known-sites dbsnp_146.hg38.vcf.gz \
        -O HCT116b_recal_data.table
```


*Way2*. Feed three bam files to `BaseRecalibrator`. Since these three bam files have the same sample ID(in this case, HCT116b), BQSR step will recoganize that they belong to the same sample.
```{r,engine='bash',eval=FALSE,way2}
gatk BaseRecalibrator \
        -R $genomeFastaFiles \
        -I SRR2481145_dedup_split.bam \
		    -I SRR2481146_dedup_split.bam \
		    -I SRR2481147_dedup_split.bam \
        --known-sites 1000G_phase1.snps.high_confidence.hg38.vcf.gz \
        --known-sites Mills_and_1000G_gold_standard.indels.hg38.vcf.gz \
		    --known-sites dbsnp_146.hg38.vcf.gz \
        -O HCT116b_recal_data.table

```

*Way3*. Perform alignment, de-duplication, indel realignment and BQSR at the replicate-level. This is then followed by joint variant calling and at the sample-level. 
```{r,eval=FALSE,engine='bash',way3}
## run HaplotypeCaller on each bam under the -ERC GVCF mode
for file in /pylon5/cc5fpcp/xiej/Cell_Line/RST2/way3/*BQSR.bam
do
time gatk HaplotypeCaller \
        -R $genomeFastaFiles \
        -I $file \
        --dont-use-soft-clipped-bases \
        -stand-call-conf 20.0 \
        -O $(basename $file _BQSR.bam).g.vcf \
		    -ERC GVCF
done

## combine gvcf
gatk CombineGVCFs \
        -R $genomeFastaFiles \
        --variant SRR2481145_1.g.vcf \
		    --variant SRR2481146_1.g.vcf \
		    --variant SRR2481147_1.g.vcf \
        -O HCT116b.g.vcf 
        
##  genoetype
gatk GenotypeGVCFs \
        -R $genomeFastaFiles \
        -V HCT116b.g.vcf \
        -O HCT116b.vcf 
```
