---
title: Optional argument in R function
author: ''
date: '2019-03-15'
slug: optional-argument-in-r-function
categories:
  - R
tags: []
---
## Background: 
I was trying to wrap several R functions into a single function, and one component function has optional argument.
```{r,eval=FALSE}
## ref is an optional argument
CLUSTERING(Raw,blocks,ref=NULL){
  ...
  if(!is.null(ref)){
    ...
  }
}
## final()is a function that will call CLUSTERING()
final <-function(i,ref=NULL){
  ...
  CLUSTERING(i,blocks,ref=NULL)
}
```
I want the function work in such a way that if provide `ref` an actual value, it returns a list of labels and a dataframe, if no `ref` is provided, just return the list of labels

The problem is that for **CLUSTERING()**,specify `ref=NULL` or provide a value to `ref`, it works as expected, while **final()** function only returns list of labels, no matter if `ref`has value or not.

After many tries&fails, I found [this post](https://dzone.com/articles/r-function-argument-lists-and-missing).

From this post, I got the idea that it is recommended to specify a value of NULL for optional argument. And it seems that something happens to the value of argument during function calling. To make clear what happens, I tried two different function calling:
```{r fcuntion calling}
f1 <- function(a='A',b=NULL) {
  print( ifelse( is.null(a), 'a not specified', paste('a =',a) ) )
  print( ifelse( is.null(b), 'b not specified', paste('b =',b) ) )
}
Callf1 <-function(a='A',b=NULL){
  print( ifelse( is.null(a), '# a not specified', paste('# a =',a) ) )
  print( ifelse( is.null(b), '# b not specified', paste('# b =',b) ) )
 f1()
}
Callf1(a='A',b=NULL)
Callf1(a='A',b='B')
Callf1_2 <-function(a='A',b=NULL){
  print( ifelse( is.null(a), '# a not specified', paste('# a =',a) ) )
  print( ifelse( is.null(b), '# b not specified', paste('# b =',b) ) )
 f1(a,b)
}
Callf1_2(a='A',b='NULL')
Callf1_2(a='A',b='B')
```

Using **Callf1()**, whether we provide value for *b* or not, it always return `b not specified`.